---
title: "SPI Journal Article - Supplmental Tables & Figures"
author: "SPI Team"
date: "`r Sys.Date()`"
output:
  bookdown::word_document2: 
    toc: yes
    fig_width: 9
    fig_height: 6
    number_sections: true

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 9, fig.height = 6, fig.cap = "")
library(tidyverse)
library(flextable)
library(here)
# devtools::install_github("worldbank/wbgviz", subdir = "wbgdata")
# devtools::install_github("worldbank/wbgviz", subdir = "wbgcharts")
# devtools::install_github("worldbank/wbgviz", subdir = "wbggeo")
# devtools::install_github("worldbank/wbgviz", subdir = "wbgmaps")
# library(wbggeo)
# library(wbgmaps)
library(ggthemes)
library(Hmisc)
library(httr)
library(patchwork)
library(ggrepel)
library(haven)
library(zoo)
library(estimatr)
library(ggpmisc)
library(ggthemes)
library(ggtext)
#set directories
dir <- here()
raw_dir <- paste(dir, '01_raw_data', sep="/")
output_dir <- paste(dir, '03_output_data', sep="/")
#weights (either unity (1) or population)
wgt <- 1
```


```{r load, include=FALSE}
#add in population
pop_df <- wbstats::wb_data(country="all",
             indicator='SP.POP.TOTL',
             start_date=2004,
             end_date=2020) %>%
  mutate(date=as.numeric(date)) %>%
  mutate(population=SP.POP.TOTL) %>%
  select(country, date, population) 
#import data
spi_index_df <- read_csv(paste(output_dir, 'SPI_index.csv', sep="/")) %>%
    left_join(pop_df) %>%
  mutate(weights=wgt)  #create a variable for weights for estimates.  This just makes is easier to switch between weights and unweighted estimates.
#read in the SPI dataset
spi_df_final <- read_csv( file = paste(output_dir, 'SPI_data.csv', sep="/")) %>%
  left_join(pop_df) %>%
  mutate(weights=wgt)  #create a variable for weights for estimates.  This just makes is easier to switch between weights and unweighted estimates.
#metadata 

metadata <- read_csv(paste(raw_dir, '/metadata/SPI_dimensions_sources.csv', sep=""))

metadata_full <- read_csv(paste(raw_dir, '/metadata/SPI_index_sources.csv', sep="")) %>%
  rename(source_name=descript) %>%
  bind_rows(metadata)

metadata <- data.frame(
  series = c('SPI.INDEX', 'SPI.INDEX.PIL1', 'SPI.INDEX.PIL2', 'SPI.INDEX.PIL3', 'SPI.INDEX.PIL4', 'SPI.INDEX.PIL5'),
  indicator_name = c('SPI Overall Score', 'Pillar 1: Data Use', 'Pillar 2: Data Services', 'Pillar 3: Data Products', 'Pillar 4: Data Sources', 'Pillar 5: Data Infrastructure')
)
#get list of country info
country_list <- wbstats::wb_countries()

SPI_WDI <- read_csv('https://raw.githubusercontent.com/worldbank/SPI/master/03_output_data/SPI_index.csv') %>%
    left_join(pop_df) %>%
  mutate(weights=wgt)  #create a variable for weights for estimates.  This just makes is easier to switch between weights and unweighted estimates.


SPI_2020 <- spi_index_df %>%
  filter(date==2020) %>%
  filter(!is.na(SPI.INDEX) & !is.na(weights)) %>%
  mutate(ISO_A3_EH=iso3c) 


SPI <- SPI_2020 %>%
  bind_rows(SPI_WDI)
#read in TopoJSON from World Bank
#countries <- geojsonio::geojson_read("WB_countries_Admin0_lowres.geojson",
#                                     what = "sp")
```

## SDG Indicator Weights Sensitivity Check

In this sensitivity check, alternative weights to the SPI overall score will be considered, where SDGs are given alternative weights.

In the preferred version of the SPI overall score, a score for each pillar is subsequently computed as the average score of the dimensions in that pillar. For pillars 1, 2, 4, and 5, the unweighted average of the dimensions within each pillar is taken.  For pillar 3 on data products, we take a weighted average of the dimensions, where the weights are based on the number of SDGs in each dimension (6 SDGs in dimension 3.1 on social statistics, 6 SDGs in dimension 3.2 on economic statistics, 2 in dimension 3.3 on environmental statistics, and 2 in dimension 3.4 on institutional statistics).   This reflects a perspective that all SDGs are of equal importance, and therefore the dimensions are weighted accordingly. Additionally, for Pillar 4 on data sources, censuses and surveys are given separate weights, so that censuses, surveys, administrative data, and geospatial data each receives a weight of 1/4. While censuses and surveys are in the same pillar in the framework, and therefore each would typically only receive a weight of 1/6 in this dimension, because of their importance in producing many indicators, they are given extra weight such that each gets a weight of 1/4. The score for each pillar (SPI.PIL_{ctp}) is calculated as follows.

In the alternative version to follow, Pillar 3 is constructed so that 1/4 weight is given to each of what we classify as Social (SDG 1-6), Economic (SDG 7-12), Environmental (SDG 13-15), and Institution (SDG 16-17).  Within each of the four groups (Social, Economic, Environmental, and Institution), the SDGs in that group are given equal weight.  For instance, the Social index score is the unweighted average of the score for SDG 1, SDG 2, SDG 3, SDG 4, SDG 5, and SDG 6.  The weights for each SDG in the preferred version of Pillar 3 shown in the body of the paper and the alternative version described above is shown in the table below.

**Table. Pillar 3 Weights for the SDGs Under Preferred and Alternative Specifications.**    

| Goal | Weight for Preferred Specification | Weight for Alternative Specification |
|---|---|---|
| SDG 1 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 2 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 3 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 4 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 5 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 6 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 7 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 8 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 9 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 10 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 11 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 12 | 1/16 | 1/24 = (1/4 * 1/6) |
| SDG 13 | 1/16 | 1/8 = (1/4 * 1/2) |
| SDG 14 | excluded | excluded |
| SDG 15 | 1/16 | 1/8 = (1/4 * 1/2) |
| SDG 16 | 1/16 | 1/8 = (1/4 * 1/2) |
| SDG 17 | 1/16 | 1/8 = (1/4 * 1/2) |




```{r index_weights_alt1}

#Pillar 1 - Overall Weight
pillar_1 <- 1/5
#Dimension 1.5: Data Use by International Organizations
dim_1_5 <- 1

#Dimension 2 - Overall Weight
pillar_2 <- 1/5
#Dimension 2.1: Data releases
dim_2_1 <- 1/3
#Dimension 2.2: Online access
dim_2_2 <- 1/3
# Dimension 2.4: Data services
dim_2_4 <- 1/3

# Dimension 3 - Overall Weight
pillar_3 <- 1/5
# Dimension 3: SDG 1
dim_3_1 <- 1/24
# Dimension 3: SDG 2
dim_3_2 <- 1/24
# Dimension 3: SDG 3
dim_3_3 <- 1/24
# Dimension 3: SDG 4
dim_3_4 <- 1/24
# Dimension 3: SDG 5
dim_3_5 <- 1/24
# Dimension 3: SDG 6
dim_3_6 <- 1/24
# Dimension 3: SDG 7
dim_3_7 <- 1/24
# Dimension 3: SDG 8
dim_3_8 <- 1/24
# Dimension 3: SDG 9
dim_3_9 <- 1/24
# Dimension 3: SDG 10
dim_3_10 <- 1/24
# Dimension 3: SDG 11
dim_3_11 <- 1/24
# Dimension 3: SDG 12
dim_3_12 <- 1/24
# Dimension 3: SDG 13
dim_3_13 <- 1/8
# Dimension 3: SDG 15
dim_3_15 <- 1/8
# Dimension 3: SDG 16
dim_3_16 <- 1/8
# Dimension 3: SDG 17
dim_3_17 <- 1/8

#Dimension 4 - Overall Weight
pillar_4 <- 1/5
# Dimension 4.1: censuses and surveys
dim_4_1.CEN <- 1/4
dim_4_1.SVY <- 1/4
#Dimension 4.2: administrative data
dim_4_2 <- 1/4
# Dimension 4.3: geospatial data
dim_4_3 <- 1/4

# Dimension 5 - Overall Weight
pillar_5 <- 1/5
# Dimension 5.2: Standards and Methods
dim_5_1 <- 0
dim_5_2 <- 1
dim_5_5 <- 0


```
  

```{r index_alt1}

# create index dataset
spi_index_df_temp <- spi_df_final %>%
  select(country, iso3c, date, starts_with("SPI"), income, region, weights, population) %>%
  arrange(-date,country)

#Drop certain indicators that don't make cut because of imcomplete coverage usually
spi_index_df_temp <- spi_index_df_temp %>%
  select( -SPI.D5.3.DISK, -SPI.D4.3.GEO.second.admin.level) 



#################
# Nested Index
################
spi_index_alt1_df <- spi_index_df_temp %>%
  arrange(country, date) %>%
  group_by(country) %>%
  mutate(across(starts_with("SPI"), na.locf, na.rm=FALSE)) %>%
  select(country, iso3c, date, everything())



#recalculate index based on the weights
pillar_total <- pillar_1 + pillar_2 + pillar_3 + pillar_4 + pillar_5
pillar2_total <- dim_2_1 + dim_2_2 + dim_2_4
pillar3_total <- 
  dim_3_1 + 
  dim_3_2 +
  dim_3_3 +
  dim_3_4 +
  dim_3_5 +
  dim_3_6 +
  dim_3_7 +
  dim_3_8 +
  dim_3_9 +
  dim_3_10 +
  dim_3_11 +
  dim_3_12 +
  dim_3_13 +
  dim_3_15 +
  dim_3_16 +
  dim_3_17
pillar4_total <- dim_4_1.CEN + dim_4_1.SVY + + dim_4_2 + dim_4_3
pillar5_total <- dim_5_1 + dim_5_2 + dim_5_5

#create index dataset
spi_index_alt1_df <- spi_index_alt1_df %>%
  mutate(SPI.DIM1.5.INDEX=rowMeans(across(starts_with("SPI.D1.5")), na.rm=FALSE),
         SPI.DIM2.1.INDEX=rowMeans(across(starts_with('SPI.D2.1'))),
         SPI.DIM2.2.INDEX=SPI.D2.2.Openness.subscore,
         SPI.DIM2.4.INDEX=SPI.D2.4.NADA,
         SPI.DIM3.1.INDEX=rowMeans(across(c("SPI.D3.1.POV",
                                          "SPI.D3.2.HNGR",
                                          "SPI.D3.3.HLTH",
                                          "SPI.D3.4.EDUC",
                                          "SPI.D3.5.GEND",
                                          "SPI.D3.6.WTRS"))),
         SPI.DIM3.2.INDEX=rowMeans(across(c("SPI.D3.7.ENRG",
                                          "SPI.D3.8.WORK",
                                          "SPI.D3.9.INDY",
                                          "SPI.D3.10.NEQL",
                                          "SPI.D3.11.CITY",
                                          "SPI.D3.12.CNSP"))),         
         SPI.DIM3.3.INDEX=rowMeans(across(c("SPI.D3.13.CLMT",
                                          "SPI.D3.15.LAND" ))),
         SPI.DIM3.4.INDEX=rowMeans(across(c("SPI.D3.16.INST",
                                          "SPI.D3.17.PTNS" ))),
         SPI.DIM4.1.CEN.INDEX=rowMeans(across(c('SPI.D4.1.1.POPU','SPI.D4.1.2.AGRI','SPI.D4.1.3.BIZZ'))), #separate census and surveys 
         SPI.DIM4.1.SVY.INDEX=rowMeans(across(c('SPI.D4.1.4.HOUS','SPI.D4.1.5.AGSVY','SPI.D4.1.6.LABR', 'SPI.D4.1.7.HLTH','SPI.D4.1.8.BZSVY'))), #separate census and surveys 
         SPI.DIM4.2.INDEX=rowMeans(across(starts_with('SPI.D4.2'))),
         SPI.DIM4.3.INDEX=rowMeans(across(starts_with('SPI.D4.3'))),
         
         SPI.DIM5.1.INDEX=rowMeans(across(starts_with('SPI.D5.1'))),
         SPI.DIM5.1.INDEX=if_else(is.na(SPI.DIM5.1.INDEX),-99,SPI.DIM5.1.INDEX),
         
         SPI.DIM5.2.INDEX=rowMeans(across(starts_with('SPI.D5.2'))),
         
         SPI.DIM5.5.INDEX=rowMeans(across(starts_with('SPI.D5.5'))),
         SPI.DIM5.5.INDEX=if_else(is.na(SPI.DIM5.5.INDEX),-99,SPI.DIM5.5.INDEX) 
             ) %>%
  mutate(
    SPI.INDEX.PIL1.ALT1=SPI.DIM1.5.INDEX,
    SPI.INDEX.PIL2.ALT1=(
      (dim_2_1/pillar2_total)*SPI.DIM2.1.INDEX +
        (dim_2_2/pillar2_total)*SPI.DIM2.2.INDEX +
        (dim_2_4/pillar2_total)*SPI.DIM2.4.INDEX )
    ,
    SPI.INDEX.PIL3.ALT1=(
      (dim_3_1/pillar3_total)*SPI.D3.1.POV +
        (dim_3_2/pillar3_total)*SPI.D3.2.HNGR +
        (dim_3_3/pillar3_total)*SPI.D3.3.HLTH +
        (dim_3_4/pillar3_total)*SPI.D3.4.EDUC +
        (dim_3_5/pillar3_total)*SPI.D3.5.GEND +
        (dim_3_6/pillar3_total)*SPI.D3.6.WTRS +
        (dim_3_7/pillar3_total)*SPI.D3.7.ENRG +
        (dim_3_8/pillar3_total)*SPI.D3.8.WORK +
        (dim_3_9/pillar3_total)*SPI.D3.9.INDY +
        (dim_3_10/pillar3_total)*SPI.D3.10.NEQL +
        (dim_3_11/pillar3_total)*SPI.D3.11.CITY +
        (dim_3_12/pillar3_total)*SPI.D3.12.CNSP +
        (dim_3_13/pillar3_total)*SPI.D3.13.CLMT +
        (dim_3_15/pillar3_total)*SPI.D3.15.LAND +
        (dim_3_16/pillar3_total)*SPI.D3.16.INST +
        (dim_3_17/pillar3_total)*SPI.D3.17.PTNS)
    ,
    SPI.INDEX.PIL4.ALT1=(
      (dim_4_1.CEN/pillar4_total)*SPI.DIM4.1.CEN.INDEX +
      (dim_4_1.SVY/pillar4_total)*SPI.DIM4.1.SVY.INDEX +
        (dim_4_2/pillar4_total)*SPI.DIM4.2.INDEX +
        (dim_4_3/pillar4_total)*SPI.DIM4.3.INDEX )
    ,
    SPI.INDEX.PIL5.ALT1=(
      (dim_5_1/pillar5_total)*SPI.DIM5.1.INDEX +
        (dim_5_2/pillar5_total)*SPI.DIM5.2.INDEX +
        (dim_5_5/pillar5_total)*SPI.DIM5.5.INDEX ),
    SPI.INDEX.ALT1=(pillar_1/pillar_total)*SPI.INDEX.PIL1.ALT1 +
      (pillar_2/pillar_total)*SPI.INDEX.PIL2.ALT1 +
      (pillar_3/pillar_total)*SPI.INDEX.PIL3.ALT1 +
      (pillar_4/pillar_total)*SPI.INDEX.PIL4.ALT1 +
      (pillar_5/pillar_total)*SPI.INDEX.PIL5.ALT1 
    #sum up based on individual dimension weights
  ) %>% #
  mutate(across(starts_with('SPI.INDEX'),~100*.)) %>%
  arrange(-date, -SPI.INDEX.ALT1) %>%
  select(country, iso3c, date, starts_with('SPI.INDEX')) 
  #  filter(date>=2016) #2016 is first year with complete data



```

```{r}

alt1_comp_df <- spi_index_alt1_df %>%
  filter(date==2020) %>%
  left_join(SPI_2020)
  

#correlations
cor_ovrl <- cor(alt1_comp_df$SPI.INDEX, alt1_comp_df$SPI.INDEX.ALT1, use='pairwise.complete' )
cor_p1 <- cor(alt1_comp_df$SPI.INDEX.PIL1, alt1_comp_df$SPI.INDEX.PIL1.ALT1, use='pairwise.complete' )
cor_p2 <- cor(alt1_comp_df$SPI.INDEX.PIL2, alt1_comp_df$SPI.INDEX.PIL2.ALT1, use='pairwise.complete' )
cor_p3 <- cor(alt1_comp_df$SPI.INDEX.PIL3, alt1_comp_df$SPI.INDEX.PIL3.ALT1, use='pairwise.complete' )
cor_p4 <- cor(alt1_comp_df$SPI.INDEX.PIL4, alt1_comp_df$SPI.INDEX.PIL4.ALT1, use='pairwise.complete' )
cor_p5 <- cor(alt1_comp_df$SPI.INDEX.PIL5, alt1_comp_df$SPI.INDEX.PIL5.ALT1, use='pairwise.complete' )
```
The correlation between the Pillar 3 score for 2020 between the preferred measure and the alternative measure is `r round(cor_p3,3)`, and the correlation between the SPI overall score under the preferred approach and the alternative approach is `r round(cor_ovrl,3)`.  The figure below shows the scatter plot between the preferred and alternative scores for Pillar 3.

```{r}

ggplot(alt1_comp_df, aes(x=SPI.INDEX.PIL3, y=SPI.INDEX.PIL3.ALT1)) +
  geom_point(alpha=0.3, color="#e76f51") +
  #geom_text_repel(aes(label=iso3c), size=3, box.padding=0.01, point.padding = 0.01, label_padding=0.01, max.overlaps = 25 ) +
  geom_text(aes(label=iso3c), color="#264653",position=position_jitter(width=.2,height=.6), check_overlap=T) +
  theme_bw() +
  xlab('SPI PIllar 3 Score - Preferred') +
  ylab('SPI PIllar 3 Score - Alternative') +
  expand_limits(x=c(0,100), y=c(0,100)) +
  labs(title = "SPI Pillar 3 Scores in 2020 for Preferred and Alternative SDG Weights",
       caption= str_wrap("Under the preferred approach, each SDG receives a weight of 1/16.  Under the alternative approach,Pillar 3 is constructed so that 1/4 weight is given to each of what we classify as Social (SDG 1-6), Economic (SDG 7-12), Environmental (SDG 13-15), and Institution (SDG 16-17).  Within each of the four groups (Social, Economic, Environmental, and Institution), the SDGs in that group are given equal weight.", 110))

```


## Pillar 4, Census and Survey  Weights Sensitivity Check

In this sensitivity check, alternative weights to the SPI overall score will be considered, where the indicator for censuses and surveys are each given weight of 1/6, rather than 1/4.

In the preferred version of the SPI overall score, a score for each pillar is subsequently computed as the average score of the dimensions in that pillar. For pillars 1, 2, 4, and 5, the unweighted average of the dimensions within each pillar is taken.  For pillar 3 on data products, we take a weighted average of the dimensions, where the weights are based on the number of SDGs in each dimension (6 SDGs in dimension 3.1 on social statistics, 6 SDGs in dimension 3.2 on economic statistics, 2 in dimension 3.3 on environmental statistics, and 2 in dimension 3.4 on institutional statistics).   This reflects a perspective that all SDGs are of equal importance, and therefore the dimensions are weighted accordingly. Additionally, for Pillar 4 on data sources, censuses and surveys are given separate weights, so that censuses, surveys, administrative data, and geospatial data each receives a weight of 1/4. While censuses and surveys are in the same pillar in the framework, and therefore each would typically only receive a weight of 1/6 in this dimension, because of their importance in producing many indicators, they are given extra weight such that each gets a weight of 1/4. The score for each pillar (SPI.PIL_{ctp}) is calculated as follows.

In the alternative version to follow, Pillar 4 is constructed so that 1/6 weight is given to censuses, 1/6 weight is given to surveys, 1/3 weight is given to administrative data, and 1/3 weight is given to geospatial data. The weights for each indicator in the preferred version of Pillar 4 shown in the body of the paper and the alternative version described above is shown in the table below.

**Table. Pillar 4 Weights for Indicators Under Preferred and Alternative Specifications.**    

| Indicator | Weight for Preferred Specification | Weight for Alternative Specification |
|---|---|---|
| Censuses | 1/4 | 1/6 = (1/2 * 1/3) |
| Surveys | 1/4 | 1/24 = (1/2 * 1/3) |
| Administrative Data | 1/4 | 1/3 |
| Geospatial Data | 1/4 | 1/3 |





```{r index_weights_alt2}

#Pillar 1 - Overall Weight
pillar_1 <- 1/5
#Dimension 1.5: Data Use by International Organizations
dim_1_5 <- 1

#Dimension 2 - Overall Weight
pillar_2 <- 1/5
#Dimension 2.1: Data releases
dim_2_1 <- 1/3
#Dimension 2.2: Online access
dim_2_2 <- 1/3
# Dimension 2.4: Data services
dim_2_4 <- 1/3

# Dimension 3 - Overall Weight
pillar_3 <- 1/5
# Dimension 3: SDG 1
dim_3_1 <- 1/24
# Dimension 3: SDG 2
dim_3_2 <- 1/24
# Dimension 3: SDG 3
dim_3_3 <- 1/24
# Dimension 3: SDG 4
dim_3_4 <- 1/24
# Dimension 3: SDG 5
dim_3_5 <- 1/24
# Dimension 3: SDG 6
dim_3_6 <- 1/24
# Dimension 3: SDG 7
dim_3_7 <- 1/24
# Dimension 3: SDG 8
dim_3_8 <- 1/24
# Dimension 3: SDG 9
dim_3_9 <- 1/24
# Dimension 3: SDG 10
dim_3_10 <- 1/24
# Dimension 3: SDG 11
dim_3_11 <- 1/24
# Dimension 3: SDG 12
dim_3_12 <- 1/24
# Dimension 3: SDG 13
dim_3_13 <- 1/8
# Dimension 3: SDG 15
dim_3_15 <- 1/8
# Dimension 3: SDG 16
dim_3_16 <- 1/8
# Dimension 3: SDG 17
dim_3_17 <- 1/8

#Dimension 4 - Overall Weight
pillar_4 <- 1/5
# Dimension 4.1: censuses and surveys
dim_4_1.CEN <- 1/6
dim_4_1.SVY <- 1/6
#Dimension 4.2: administrative data
dim_4_2 <- 1/3
# Dimension 4.3: geospatial data
dim_4_3 <- 1/3

# Dimension 5 - Overall Weight
pillar_5 <- 1/5
# Dimension 5.2: Standards and Methods
dim_5_1 <- 0
dim_5_2 <- 1
dim_5_5 <- 0


```
  

```{r index_alt2}

# create index dataset
spi_index_df_temp <- spi_df_final %>%
  select(country, iso3c, date, starts_with("SPI"), income, region, weights, population) %>%
  arrange(-date,country)

#Drop certain indicators that don't make cut because of imcomplete coverage usually
spi_index_df_temp <- spi_index_df_temp %>%
  select( -SPI.D5.3.DISK, -SPI.D4.3.GEO.second.admin.level) 



#################
# Nested Index
################
spi_index_alt1_df <- spi_index_df_temp %>%
  arrange(country, date) %>%
  group_by(country) %>%
  mutate(across(starts_with("SPI"), na.locf, na.rm=FALSE)) %>%
  select(country, iso3c, date, everything())



#recalculate index based on the weights
pillar_total <- pillar_1 + pillar_2 + pillar_3 + pillar_4 + pillar_5
pillar2_total <- dim_2_1 + dim_2_2 + dim_2_4
pillar3_total <- 
  dim_3_1 + 
  dim_3_2 +
  dim_3_3 +
  dim_3_4 +
  dim_3_5 +
  dim_3_6 +
  dim_3_7 +
  dim_3_8 +
  dim_3_9 +
  dim_3_10 +
  dim_3_11 +
  dim_3_12 +
  dim_3_13 +
  dim_3_15 +
  dim_3_16 +
  dim_3_17
pillar4_total <- dim_4_1.CEN + dim_4_1.SVY + + dim_4_2 + dim_4_3
pillar5_total <- dim_5_1 + dim_5_2 + dim_5_5

#create index dataset
spi_index_alt2_df <- spi_index_alt1_df %>%
  mutate(SPI.DIM1.5.INDEX=rowMeans(across(starts_with("SPI.D1.5")), na.rm=FALSE),
         SPI.DIM2.1.INDEX=rowMeans(across(starts_with('SPI.D2.1'))),
         SPI.DIM2.2.INDEX=SPI.D2.2.Openness.subscore,
         SPI.DIM2.4.INDEX=SPI.D2.4.NADA,
         SPI.DIM3.1.INDEX=rowMeans(across(c("SPI.D3.1.POV",
                                          "SPI.D3.2.HNGR",
                                          "SPI.D3.3.HLTH",
                                          "SPI.D3.4.EDUC",
                                          "SPI.D3.5.GEND",
                                          "SPI.D3.6.WTRS"))),
         SPI.DIM3.2.INDEX=rowMeans(across(c("SPI.D3.7.ENRG",
                                          "SPI.D3.8.WORK",
                                          "SPI.D3.9.INDY",
                                          "SPI.D3.10.NEQL",
                                          "SPI.D3.11.CITY",
                                          "SPI.D3.12.CNSP"))),         
         SPI.DIM3.3.INDEX=rowMeans(across(c("SPI.D3.13.CLMT",
                                          "SPI.D3.15.LAND" ))),
         SPI.DIM3.4.INDEX=rowMeans(across(c("SPI.D3.16.INST",
                                          "SPI.D3.17.PTNS" ))),
         SPI.DIM4.1.CEN.INDEX=rowMeans(across(c('SPI.D4.1.1.POPU','SPI.D4.1.2.AGRI','SPI.D4.1.3.BIZZ'))), #separate census and surveys 
         SPI.DIM4.1.SVY.INDEX=rowMeans(across(c('SPI.D4.1.4.HOUS','SPI.D4.1.5.AGSVY','SPI.D4.1.6.LABR', 'SPI.D4.1.7.HLTH','SPI.D4.1.8.BZSVY'))), #separate census and surveys 
         SPI.DIM4.2.INDEX=rowMeans(across(starts_with('SPI.D4.2'))),
         SPI.DIM4.3.INDEX=rowMeans(across(starts_with('SPI.D4.3'))),
         
         SPI.DIM5.1.INDEX=rowMeans(across(starts_with('SPI.D5.1'))),
         SPI.DIM5.1.INDEX=if_else(is.na(SPI.DIM5.1.INDEX),-99,SPI.DIM5.1.INDEX),
         
         SPI.DIM5.2.INDEX=rowMeans(across(starts_with('SPI.D5.2'))),
         
         SPI.DIM5.5.INDEX=rowMeans(across(starts_with('SPI.D5.5'))),
         SPI.DIM5.5.INDEX=if_else(is.na(SPI.DIM5.5.INDEX),-99,SPI.DIM5.5.INDEX) 
             ) %>%
  mutate(
    SPI.INDEX.PIL1.ALT2=SPI.DIM1.5.INDEX,
    SPI.INDEX.PIL2.ALT2=(
      (dim_2_1/pillar2_total)*SPI.DIM2.1.INDEX +
        (dim_2_2/pillar2_total)*SPI.DIM2.2.INDEX +
        (dim_2_4/pillar2_total)*SPI.DIM2.4.INDEX )
    ,
    SPI.INDEX.PIL3.ALT2=(
      (dim_3_1/pillar3_total)*SPI.D3.1.POV +
        (dim_3_2/pillar3_total)*SPI.D3.2.HNGR +
        (dim_3_3/pillar3_total)*SPI.D3.3.HLTH +
        (dim_3_4/pillar3_total)*SPI.D3.4.EDUC +
        (dim_3_5/pillar3_total)*SPI.D3.5.GEND +
        (dim_3_6/pillar3_total)*SPI.D3.6.WTRS +
        (dim_3_7/pillar3_total)*SPI.D3.7.ENRG +
        (dim_3_8/pillar3_total)*SPI.D3.8.WORK +
        (dim_3_9/pillar3_total)*SPI.D3.9.INDY +
        (dim_3_10/pillar3_total)*SPI.D3.10.NEQL +
        (dim_3_11/pillar3_total)*SPI.D3.11.CITY +
        (dim_3_12/pillar3_total)*SPI.D3.12.CNSP +
        (dim_3_13/pillar3_total)*SPI.D3.13.CLMT +
        (dim_3_15/pillar3_total)*SPI.D3.15.LAND +
        (dim_3_16/pillar3_total)*SPI.D3.16.INST +
        (dim_3_17/pillar3_total)*SPI.D3.17.PTNS)
    ,
    SPI.INDEX.PIL4.ALT2=(
      (dim_4_1.CEN/pillar4_total)*SPI.DIM4.1.CEN.INDEX +
      (dim_4_1.SVY/pillar4_total)*SPI.DIM4.1.SVY.INDEX +
        (dim_4_2/pillar4_total)*SPI.DIM4.2.INDEX +
        (dim_4_3/pillar4_total)*SPI.DIM4.3.INDEX )
    ,
    SPI.INDEX.PIL5.ALT2=(
      (dim_5_1/pillar5_total)*SPI.DIM5.1.INDEX +
        (dim_5_2/pillar5_total)*SPI.DIM5.2.INDEX +
        (dim_5_5/pillar5_total)*SPI.DIM5.5.INDEX ),
    SPI.INDEX.ALT2=(pillar_1/pillar_total)*SPI.INDEX.PIL1.ALT2 +
      (pillar_2/pillar_total)*SPI.INDEX.PIL2.ALT2 +
      (pillar_3/pillar_total)*SPI.INDEX.PIL3.ALT2 +
      (pillar_4/pillar_total)*SPI.INDEX.PIL4.ALT2 +
      (pillar_5/pillar_total)*SPI.INDEX.PIL5.ALT2 
    #sum up based on individual dimension weights
  ) %>% #
  mutate(across(starts_with('SPI.INDEX'),~100*.)) %>%
  arrange(-date, -SPI.INDEX.ALT2) %>%
  select(country, iso3c, date, starts_with('SPI.INDEX')) 
  #  filter(date>=2016) #2016 is first year with complete data



```

```{r}

alt2_comp_df <- spi_index_alt2_df %>%
  filter(date==2020) %>%
  left_join(SPI_2020)
  

#correlations
cor_ovrl <- cor(alt2_comp_df$SPI.INDEX, alt2_comp_df$SPI.INDEX.ALT2, use='pairwise.complete' )
cor_p1 <- cor(alt2_comp_df$SPI.INDEX.PIL1, alt2_comp_df$SPI.INDEX.PIL1.ALT2, use='pairwise.complete' )
cor_p2 <- cor(alt2_comp_df$SPI.INDEX.PIL2, alt2_comp_df$SPI.INDEX.PIL2.ALT2, use='pairwise.complete' )
cor_p3 <- cor(alt2_comp_df$SPI.INDEX.PIL3, alt2_comp_df$SPI.INDEX.PIL3.ALT2, use='pairwise.complete' )
cor_p4 <- cor(alt2_comp_df$SPI.INDEX.PIL4, alt2_comp_df$SPI.INDEX.PIL4.ALT2, use='pairwise.complete' )
cor_p5 <- cor(alt2_comp_df$SPI.INDEX.PIL5, alt2_comp_df$SPI.INDEX.PIL5.ALT2, use='pairwise.complete' )
```
The correlation between the Pillar 4 score for 2020 between the preferred measure and the alternative measure is `r round(cor_p3,3)`, and the correlation between the SPI overall score under the preferred approach and the alternative approach is `r round(cor_ovrl,3)`.  The figure below shows the scatter plot between the preferred and alternative scores for Pillar 4.

```{r}

ggplot(alt2_comp_df, aes(x=SPI.INDEX.PIL4, y=SPI.INDEX.PIL4.ALT2)) +
  geom_point(alpha=0.3, color="#e76f51") +
  #geom_text_repel(aes(label=iso3c), size=3, box.padding=0.01, point.padding = 0.01, label_padding=0.01, max.overlaps = 25 ) +
  geom_text(aes(label=iso3c), color="#264653",position=position_jitter(width=.2,height=.6), check_overlap=T) +
  theme_bw() +
  xlab('SPI PIllar 4 Score - Preferred') +
  ylab('SPI PIllar 4 Score - Alternative') +
  expand_limits(x=c(0,100), y=c(0,100)) +
  labs(title = "SPI Pillar 4 Scores in 2020 for Preferred and Alternative Indicator Weights",
       caption= str_wrap("Under the preferred approach, each indicator (censuses, surveys, administrative, and geospatial) receive weights of 1/4.  In the alternative version to follow, Pillar 4 is constructed so that 1/6 weight is given to censuses, 1/6 weight is given to surveys, 1/3 weight is given to administrative data, and 1/3 weight is given to geospatial data.", 110))

```


## Additional Findings by Pillar

The tables and charts below highlight differences across countries by income group in specific indicators.

```{r}

spi_income_table <- function(data, indicators, reference_year) {

      df_overall <- get(data) %>%
      filter(date==as.numeric(reference_year)) %>% 
      select(country, iso3c, date, income, region, all_of(indicators), SPI.INDEX) 
    
    
    spi_groups_quantiles <- quantile(df_overall$SPI.INDEX, probs=c(1,2,3,4)/5,na.rm=T)
    
    df_overall <- df_overall %>%
      mutate(spi_groups=case_when(
        between(SPI.INDEX, spi_groups_quantiles[4],100) ~ "Top Quintile",
        between(SPI.INDEX, spi_groups_quantiles[3],spi_groups_quantiles[4]) ~ "4th Quintile",
        between(SPI.INDEX, spi_groups_quantiles[2],spi_groups_quantiles[3]) ~ "3rd Quintile",
        between(SPI.INDEX, spi_groups_quantiles[1],spi_groups_quantiles[2]) ~ "2nd Quintile",
        between(SPI.INDEX, 0,spi_groups_quantiles[1]) ~ "Bottom 20%"
      )) %>%
      mutate(spi_groups=factor(spi_groups, 
                               levels=c("Top Quintile","4th Quintile","3rd Quintile","2nd Quintile","Bottom 20%" )))  
    
    #produce by income
    sumstats<- df_overall %>%
      group_by(income) %>%
      filter(income %in% c("Low income", "Lower middle income", "Upper middle income", "High income" )) %>%
      select(income, all_of(indicators)) %>%
      summarise_all(~round(mean(., na.rm=T),1)) 
    
    #produce global number
    sumstats_gl<- df_overall %>%
      mutate(income='Global') %>%
      group_by(income) %>%
      select(income, all_of(indicators)) %>%
      summarise_all(~round(mean(., na.rm=T),1)) 
    
    
    #transpose data
    sumstats_df_long <-sumstats 
    
    sumstats_df <- as.data.frame(t(sumstats_df_long %>% select(-income)))
    colnames(sumstats_df) = sumstats_df_long$income 
    
    
    sumstats_df <- sumstats_df %>%
      rownames_to_column() %>%
      rename(series=rowname)
    
    
    #create labels df
    metadata_tab2_overall <- metadata_full %>% 
      janitor::clean_names() %>%
      transmute(series=source_id, 
                indicator_name=source_name,
                pillar=pillar)
    
    
    #add variable label
    sumstats_df <- sumstats_df %>%
      left_join(metadata_tab2_overall) %>%
      rename(Pillar=pillar,
             Series=series,
             Label=indicator_name) %>%
      mutate(Label=if_else(is.na(Label),Series,Label)) %>%
      select(Pillar, Label, c("Low income", "Lower middle income", "Upper middle income", "High income" ))

      sumstats_df
 

}

FitFlextableToPage <- function(ft, pgwidth = 6){

  ft_out <- ft %>% 
    add_footer_lines(values = "SPI index scored on 0-100 scale.  Indicators scored on 0-1 scale. \n Source: World Bank. Statistical Performance Indicators."                 ) %>%
    autofit()

  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}

```

### Selected Indicators

**Table. Select SPI Indicators by Income Group in 2020 **
```{r sel}
indicators<-c("SPI.D1.5.DT.TDS.DPPF.XP.ZS","SPI.D2.2.Openness.subscore","SPI.D3.1.POV","SPI.D3.10.NEQL",
             "SPI.D4.1.1.POPU", "SPI.D4.2.3.CRVS", "SPI.D4.3.GEO.first.admin.level","SPI.D5.2.4.CPIBY", "SPI.D5.2.6.EMPL")

income_tab <- spi_income_table('spi_index_df',indicators,2020) 

tib <- as_grouped_data(x=income_tab, groups=c("Pillar"))

tib %>%
  as_flextable( ) %>% 
  flextable::compose(
    i = ~ !is.na(Pillar), # when var_group not NA
    j = "Label", # on column "var"
    # create a paragraph containing a chunk containing value of `var_group`
    value = as_paragraph(as_chunk(Pillar, props=officer::fp_text(bold=TRUE, color="#48cae4",font.size=12)))) %>% 
  hline(i = ~ !is.na(Pillar), border = officer::fp_border() ) %>% 
  autofit() %>%
  add_footer_lines(values = "SPI index scored on 0-100 scale.  Indicators scored on 0-1 scale. \n Source: World Bank. Statistical Performance Indicators."                 ) %>%
  fontsize(i=1,size=12, part='header') 


```


### Pillar 1: Data Use

**Table. Pillar 1: Data Use SPI Indicators by Income Group in 2020 **
```{r p1}

indicators<-c("SPI.D1.5.POV","SPI.D1.5.CHLD.MORT","SPI.D1.5.DT.TDS.DPPF.XP.ZS", "SPI.D1.5.SAFE.MAN.WATER", "SPI.D1.5.LFP")

income_tab <- spi_income_table('spi_index_df',indicators,2020) 

tib <- as_grouped_data(x=income_tab, groups=c("Pillar"))

tib %>%
  as_flextable( ) %>% 
  flextable::compose(
    i = ~ !is.na(Pillar), # when var_group not NA
    j = "Label", # on column "var"
    # create a paragraph containing a chunk containing value of `var_group`
    value = as_paragraph(as_chunk(Pillar, props=officer::fp_text(bold=TRUE, color="#48cae4",font.size=12)))) %>% 
  hline(i = ~ !is.na(Pillar), border = officer::fp_border() ) %>% 
  autofit() %>%
  add_footer_lines(values = "SPI index scored on 0-100 scale.  Indicators scored on 0-1 scale. \n Source: World Bank. Statistical Performance Indicators."                 ) %>%
  fontsize(i=1,size=12, part='header') 

```

### Pillar 2: Data Services

**Table. Pillar 2: Data Services SPI Indicators by Income Group in 2020 **
```{r p2}

indicators<-c("SPI.D2.1.GDDS",
              "SPI.D2.2.Openness.subscore",
              "SPI.D2.2.Machine.readable",      "SPI.D2.2.Non.proprietary",
              "SPI.D2.2.Download.options",      "SPI.D2.2.Metadata.available",
              "SPI.D2.2.Terms.of.use",
              "SPI.D2.4.NADA")

income_tab <- spi_income_table('spi_index_df',indicators,2020) 

tib <- as_grouped_data(x=income_tab, groups=c("Pillar"))

tib %>%
  as_flextable( ) %>% 
  flextable::compose(
    i = ~ !is.na(Pillar), # when var_group not NA
    j = "Label", # on column "var"
    # create a paragraph containing a chunk containing value of `var_group`
    value = as_paragraph(as_chunk(Pillar, props=officer::fp_text(bold=TRUE, color="#48cae4",font.size=12)))) %>% 
  hline(i = ~ !is.na(Pillar), border = officer::fp_border() ) %>% 
  autofit() %>%
  add_footer_lines(values = "SPI index scored on 0-100 scale.  Indicators scored on 0-1 scale. \n Source: World Bank. Statistical Performance Indicators."                 ) %>%
  fontsize(i=1,size=12, part='header') 

```

### Pillar 3: Data Products

**Table. Pillar 3: Data Products SPI Indicators by Income Group in 2020 **
```{r p3}

indicators<-c("SPI.D3.1.POV",
              "SPI.D3.2.HNGR",
              "SPI.D3.3.HLTH",
              "SPI.D3.4.EDUC",
              "SPI.D3.5.GEND",
              "SPI.D3.6.WTRS",
              "SPI.D3.7.ENRG",
              "SPI.D3.8.WORK",
              "SPI.D3.9.INDY",
              "SPI.D3.10.NEQL",
              "SPI.D3.11.CITY",
              "SPI.D3.12.CNSP",
              "SPI.D3.13.CLMT",
              "SPI.D3.15.LAND",
              "SPI.D3.16.INST",
              "SPI.D3.17.PTNS")

income_tab <- spi_income_table('spi_index_df',indicators,2020) 

tib <- as_grouped_data(x=income_tab, groups=c("Pillar"))

tib %>%
  as_flextable( ) %>% 
  flextable::compose(
    i = ~ !is.na(Pillar), # when var_group not NA
    j = "Label", # on column "var"
    # create a paragraph containing a chunk containing value of `var_group`
    value = as_paragraph(as_chunk(Pillar, props=officer::fp_text(bold=TRUE, color="#48cae4",font.size=12)))) %>% 
  hline(i = ~ !is.na(Pillar), border = officer::fp_border() ) %>% 
  autofit() %>%
  add_footer_lines(values = "SPI index scored on 0-100 scale.  Indicators scored on 0-1 scale. \n Source: World Bank. Statistical Performance Indicators."                 ) %>%
  fontsize(i=1,size=12, part='header') 

```

### Pillar 4: Data Sources

**Table. Pillar 4: Data Sources SPI Indicators by Income Group in 2020 **
```{r p4}

indicators<-c(
              "SPI.D4.1.1.POPU",
              "SPI.D4.1.2.AGRI",
              "SPI.D4.1.3.BIZZ",
              "SPI.D4.1.4.HOUS",
              "SPI.D4.1.5.AGSVY",
              "SPI.D4.1.6.LABR",
              "SPI.D4.1.7.HLTH",
              "SPI.D4.1.8.BZSVY",
              "SPI.D4.2.3.CRVS",
              "SPI.D4.3.GEO.first.admin.level")

income_tab <- spi_income_table('spi_index_df',indicators,2020) 

tib <- as_grouped_data(x=income_tab, groups=c("Pillar"))

tib %>%
  as_flextable( ) %>% 
  flextable::compose(
    i = ~ !is.na(Pillar), # when var_group not NA
    j = "Label", # on column "var"
    # create a paragraph containing a chunk containing value of `var_group`
    value = as_paragraph(as_chunk(Pillar, props=officer::fp_text(bold=TRUE, color="#48cae4",font.size=12)))) %>% 
  hline(i = ~ !is.na(Pillar), border = officer::fp_border() ) %>% 
  autofit() %>%
  add_footer_lines(values = "SPI index scored on 0-100 scale.  Indicators scored on 0-1 scale. \n Source: World Bank. Statistical Performance Indicators."                 ) %>%
  fontsize(i=1,size=12, part='header') 

```

### Pillar 5: Data Infrastructure

**Table. Pillar 5: Data Infrastructure SPI Indicators by Income Group in 2020 **
```{r p5}

indicators<-c(
              #"SPI.D5.1.DILG",
              "SPI.D5.2.1.SNAU",
              "SPI.D5.2.2.NABY",
              "SPI.D5.2.3.CNIN",
              "SPI.D5.2.4.CPIBY",
              "SPI.D5.2.5.HOUS",
              "SPI.D5.2.6.EMPL",
              "SPI.D5.2.7.CGOV",
              "SPI.D5.2.8.FINA",
              "SPI.D5.2.9.MONY",
              "SPI.D5.2.10.GSBP")
              #"SPI.D5.5.DIFI")

income_tab <- spi_income_table('spi_index_df',indicators,2020) 

tib <- as_grouped_data(x=income_tab, groups=c("Pillar"))

tib %>%
  as_flextable( ) %>% 
  flextable::compose(
    i = ~ !is.na(Pillar), # when var_group not NA
    j = "Label", # on column "var"
    # create a paragraph containing a chunk containing value of `var_group`
    value = as_paragraph(as_chunk(Pillar, props=officer::fp_text(bold=TRUE, color="#48cae4",font.size=12)))) %>% 
  hline(i = ~ !is.na(Pillar), border = officer::fp_border() ) %>% 
  autofit() %>%
  add_footer_lines(values = "SPI index scored on 0-100 scale.  Indicators scored on 0-1 scale. \n Source: World Bank. Statistical Performance Indicators."                 ) %>%
  fontsize(i=1,size=12, part='header') 
```

## Recency of Census

**Figure. Date of Most Recent Population and Housing Census by Fragile and Conflict Situation Status**
```{r, fig.width=15, fig.height=14}

    #FY21 Fragile and conflict-affected situations (http://pubdocs.worldbank.org/en/888211594267968803/FCSList-FY21.pdf)
  high_intensity_conflict <- c('Afghanistan', 'Libya', 'Somalia', 'Syrian Arab Republic' )
  
  medium_intensity_conflict <- c('Burkina Faso', 'Cameroon','Central African Republic', 'Chad', 'Congo, Dem. Rep.',
                                 'Iraq','Mali','Mozambique','Myanmar','Niger','Nigeria','South Sudan','Yemen, Rep.')
  
  high_institutional_social_fragility <- c('Burundi','Congo, Rep.','Eritrea','Gambia, The','Guinea-Bissau',
                                           'Haiti','Kosovo','Lao PDR','Lebanon','Liberia','Papua New Guinea',
                                           'Sudan','Venezuela, RB','West Bank and Gaza','Zimbabwe')
  
  small_states <- c('Comoros','Kiribati','Marshall Islands','Micronesia, Fed. Sts.','Solomon Islands','Timor-Leste','Tuvalu')

census_age_df <- spi_df_final %>%
  select(country, iso3c, date, income, region, SPI.D4.1.1.POPU, RAW.D4.1.1.POPU) %>%
  mutate(pop_census_age=as.numeric(str_extract(RAW.D4.1.1.POPU, "\\d{4}"))) %>%
  filter(date==2020) %>%
  mutate(fcs=case_when( #create indicators for Fragile and Conflict-affected Situations
      country %in% high_intensity_conflict ~ "FCS country",
      country %in% medium_intensity_conflict ~ "FCS country",
      country %in% high_institutional_social_fragility ~ "FCS country",
      country %in% small_states ~ "FCS country",
      TRUE ~ "Non-FCS country"))
  
census_age_df <- census_age_df %>%
  group_by(pop_census_age) %>%
  mutate(country_len = row_number() )

ggplot(census_age_df, aes(x=pop_census_age, y=country_len, color=fcs)) +
  geom_text(aes(label=iso3c), size=4,position=position_jitter(width=.2,height=.6), check_overlap = TRUE) +
  theme_bw() +
  xlab('Year of Last Population Census') +
  ylab('Country') +
  expand_limits(x=c(1940:2025)) +
  theme(
    axis.title = element_text(size=18),
    axis.text = element_text(size=15),
    legend.text=element_text(size=15),
  
    legend.position = 'bottom'
  )

```

## Recency of National Account Base Year

**Figure. Date of National Account Base Year by Income**
```{r naby, fig.width=15, fig.height=14}

naby_age_df <- spi_df_final %>%
  select(country, iso3c, date, income, region, SPI.D5.2.2.NABY, RAW.D5.2.2.NABY) %>%
  mutate(naby_age=case_when(
    RAW.D5.2.2.NABY=="Original chained constant price data are rescaled." ~ 2020, 
    TRUE ~  as.numeric(RAW.D5.2.2.NABY))
  ) %>%
  filter(date==2020) %>%
  mutate(fcs=case_when( #create indicators for Fragile and Conflict-affected Situations
      country %in% high_intensity_conflict ~ "FCS country",
      country %in% medium_intensity_conflict ~ "FCS country",
      country %in% high_institutional_social_fragility ~ "FCS country",
      country %in% small_states ~ "FCS country",
      TRUE ~ "Non-FCS country")) %>%
  filter(country!="Puerto Rico") %>%
  filter(income!="Not classified")
  
naby_age_df <- naby_age_df %>%
  group_by(naby_age) %>%
  mutate(country_len = row_number() )

pal <- c("High income"= "#264653", "Upper middle income" = "#2a9d8f","Lower middle income"= "#f4a261","Low income"= "#e76f51")

ggplot(naby_age_df, aes(x=naby_age, y=country_len, color=income)) +
  geom_text(aes(label=iso3c), size=4,position=position_jitter(width=.2,height=.6), check_overlap = TRUE) +
  theme_bw() +
  xlab('Year of National Accounts Base Year') +
  ylab('Country') +
  scale_color_manual(
    values=pal
  ) +
  expand_limits(x=c(1980:2025)) +
  theme(
    axis.title = element_text(size=18),
    axis.text = element_text(size=15),
    legend.text=element_text(size=15),
    legend.position = 'bottom'
  )


```

